---
title: "Climate Change"
format: 
  revealjs:
    slide-number: true
---

Constructor University, Data Science for Business and Society

Name: Sagar Kumar and Betül Yurtman

Matriculation No.: 30000402 and 30008426

Course: MDSSBO-203-Visual Communication and Data Storytelling

Semester: Fall 2025

Instructor: Prof. Jan Lorenz

## Climate Change

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(knitr)
library(scales)
library(readxl)
library(ggthemes)
```

## Dataset Sources

https://ourworldindata.org/co2-emissions : owid-co2-data.xlsx & annual-co2-emissions-per-country.csv

https://github.com/owid/owid-datasets/ : Global average temperature anomaly - Hadley Centre.csv

https://data.worldbank.org/ : Country-Year-Population.xlsx

## Insight 1: Global Temperature is Rising

## Climate Stripes

```{r}
# Dataset
temp_raw <- read_csv("datasets/Global average temperature anomaly - Hadley Centre.csv")

# Keep median anomaly only
temp_med <- temp_raw %>%
  filter(Entity == "median") %>%
  rename(year = Year, anomaly = `Global average temperature anomaly (Hadley Centre)`)

# Values for the title and the subtitle
start_year <- min(temp_med$year, na.rm = TRUE)
end_year <- max(temp_med$year, na.rm = TRUE)
temp_min <- round(min(temp_med$anomaly, na.rm = TRUE), 2)
temp_max <- round(max(temp_med$anomaly, na.rm = TRUE), 2)

# Plot 
p_stripes <- ggplot(temp_med, aes(x = year, y = 1, fill = anomaly)) +
  geom_tile() +
  scale_fill_gradient2(
  low = "#08306B",     
  mid = "#FFFFFF", 
  high = "#B10026",  
  midpoint = 0,
  guide = "none"
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    breaks = seq(1900, max(temp_med$year), by = 10)  # Year labels every 10 years
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    title = paste0("Global Climate Stripes (", start_year, "–", end_year, ")"),
    subtitle = paste0(
      "Median global temperature (",
      temp_min, "°C to ", temp_max, "°C)"
    ),
    x = "Year",
    y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  )

p_stripes
```

This climate-stripes chart visualizes how global temperatures have changed from 1948 to 2017. Each vertical stripe represents one year, with blue tones indicating cooler-than-average years and red tones indicating warmer-than-average years. The progression from predominantly blue stripes in the mid-20th century to deep red stripes in recent decades clearly illustrates the long-term warming trend. This simple but powerful graphic highlights how global temperatures have steadily increased over time, reflecting the accelerating impact of human-driven climate change.

```{r}
# Reading the data
data <- read_excel("datasets/owid-co2-data.xlsx", sheet = "Data")
glimpse(data)
head(data)
```

## Global CO₂ Emissions and Their Contribution to Global Temperature Change (1950–Present)

```{r}
temp_dat <- data %>%
  filter(!is.na(temperature_change_from_co2), !is.na(co2), year >= 1950)

temp_world <- temp_dat %>%
  filter(country == "World", year >= 1950) %>%
  select(year, global_co2 = co2, global_temp_effect = temperature_change_from_co2) %>%
  filter(!is.na(global_co2), !is.na(global_temp_effect))

p_temp <- ggplot(temp_world, aes(global_co2, global_temp_effect)) +
  geom_line(color = "#C4463A", linewidth = 1.2) +
  geom_point(size = 2) +
  scale_x_continuous(labels = label_number(big.mark = ",")) +
  labs(
    title = "Global CO2 Emissions and Contribution to Temperature Change (1950–Present)",
    subtitle = "World series from OWID; higher emissions align with greater warming contribution",
    x = "Global CO2 emissions (million tonnes)",
    y = "Temperature change attributed to CO2 (°C)"
  ) +
  theme_minimal(base_size = 13)

p_temp
```

This chart shows how global carbon dioxide (CO₂) emissions have risen since 1950 and how this rise corresponds to an increasing contribution to global temperature change. As the line moves upward and to the right, it illustrates that higher annual global CO₂ emissions are consistently associated with greater warming influence on the Earth's climate. The relationship is not random—temperature impact steadily increases as emissions rise, reflecting the cumulative nature of CO₂ in the atmosphere. This visualization highlights the close connection between human-driven emissions and global warming, emphasizing that continued emission increases lead to progressively stronger climate effects.

## Insight 2: Economic Growth Shapes CO₂ Emissions

## GDP per capita vs. CO2 per capita

```{r}
# Build a compact dataset
df_simple <- data %>%
  select(country, year, population, gdp, co2) %>%     # keep only essentials
  filter(year >= 1950) %>%                            # modern period
  filter(!is.na(population), !is.na(gdp), !is.na(co2),
         population > 0) %>%                          # valid rows only
  mutate(
    gdp_pc = gdp / population,                        # GDP per capita
    co2_pc = co2 / population                         # CO2 per capita
  )

# Use the latest year that actually has both per-capita metrics
latest <- df_simple %>%
  filter(!is.na(gdp_pc), !is.na(co2_pc)) %>%
  summarise(latest = max(year, na.rm = TRUE)) %>%
  pull(latest)

# Slice the data for that year
df_latest <- df_simple %>%
  filter(year == latest)

# Countries to label on the plot
label_countries <- c(
  "United States", "China", "India",
  "Germany", "Japan", "Turkey",
  "Brazil", "Russia", "United Kingdom"
)

# Scatter plot: GDP per capita vs CO2 per capita (log-log)
# LOESS shows the general trend without assuming linearity.
p_scatter <- ggplot(df_latest, aes(x = gdp_pc, y = co2_pc)) +
  geom_point(alpha = 0.6, color = "grey40") +                  # all countries
  # Labels for selected countries
  geom_text(
    data = df_latest %>% filter(country %in% label_countries),
    aes(label = country),
    size = 3,
    nudge_x = 0.03,
    nudge_y = 0.03,
    check_overlap = TRUE
  ) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 1) +   # trend line
  scale_x_log10(labels = label_number(scale_cut = cut_short_scale())) +
  scale_y_log10(labels = label_number(scale_cut = cut_short_scale())) +
  labs(
    title = paste("Relationship between GDP per capita and CO2 per capita —", latest),
    subtitle = "Higher income levels are generally associated with higher CO2 emissions per person.",
    x = "GDP per capita (log scale)",
    y = "CO2 emissions per capita (log scale)"
  ) +
  theme_minimal(base_size = 14)

p_scatter
```

This chart shows that countries with higher income levels tend to emit more CO₂ per person. Each point represents a country, and both axes use a log scale to allow comparison across very different economies. The upward trend indicates that as GDP per capita increases, CO₂ emissions per capita also generally rise. However, the points are spread out rather than forming a perfect line, which means income alone does not determine emissions. Some wealthy countries emit less because they use cleaner energy sources or have strong climate policies, while some rapidly developing countries emit more due to coal-based industrial growth.

## Economic Development and CO₂ Emissions Pathways (1990–Present)

```{r}
# Countries to compare
countries <- c("Turkey", "United States", "China", "Germany", "India", "Brazil", "Russia")

# Filter: keep valid vars and strictly positive values for log axes
df_paths <- df_simple %>%
  filter(country %in% countries,
         year >= 1990,
         !is.na(gdp_pc), !is.na(co2_pc),
         gdp_pc > 0, co2_pc > 0)

# Latest point per country to highlight
latest_points <- df_paths %>%
  group_by(country) %>%
  slice_max(year, with_ties = FALSE) %>%
  ungroup()

# Dynamic years for subtitle
end_year_paths   <- max(df_paths$year, na.rm = TRUE)

p_paths <- ggplot(df_paths, aes(x = gdp_pc, y = co2_pc, color = country)) +
  geom_path(linewidth = 1, alpha = 0.9,
            arrow = arrow(type = "closed", length = unit(0.12, "in"))) +
  geom_point(data = latest_points, size = 3) +

  scale_x_log10(labels = label_number(big.mark = ",")) +
  scale_y_log10(labels = label_number(big.mark = ",")) +
  labs(
    title    = "GDP per capita → CO2 per capita Pathways (1990–Present)",
    subtitle = paste0("Arrows indicate direction over time (1990 → ", end_year_paths, ")"
    ),
    x = "GDP per capita (log scale)",
    y = "CO2 emissions per capita (log scale)",
    color = "Country"
  ) +
  theme_minimal(base_size = 14)

p_paths

```

This figure shows how economic development relates to CO₂ emissions over time for selected countries (1990 → present). Each line represents a country’s trajectory: as GDP per capita increases (moving right), CO₂ emissions per capita may increase or decrease (moving up or down). The arrows indicate the direction of change over time.

Country & Interpretation China & India: Rapid economic growth accompanied by rising emissions, reflecting carbon-intensive industrialization.

Turkey & Brazil: Moderate income growth with increasing emissions, indicating continued dependence on fossil fuels.

Russia: High emissions per capita with relatively unstable economic growth, shaped by energy-heavy industries and resource extraction.

United States: Very high income with declining emissions per person, suggesting partial decoupling through cleaner energy and efficiency.

Germany: High income with clear, long-term decline in emissions, showing successful low-carbon transition policies.

Countries do not follow the same development–emissions pathway. Some countries (e.g., Germany, United States) reduce emissions while maintaining high income, while others (e.g., China, Turkey) still experience rising emissions as their economies grow.

## Insight 3: Climate responsibility is unequal - historical emissions are concentrated in rich countries

## Question 1: Which Countries Emit The Most Carbon Today?

```{r}
# read data
annual_emmisions_data <- read_csv("datasets/annual-co2-emissions-per-country.csv", show_col_types = FALSE)

# exclude known regions or groups
regions <- c(
  "Asia", "Africa", "North America", "South America", "Oceania",
  "European Union (27)", "European Union (28)", "Middle East (GCP)",
  "High-income countries", "Low-income countries", "Upper-middle-income countries",
  "Lower-middle-income countries", "Non-OECD (GCP)", "OECD (GCP)", "Asia (GCP)",
  "Europe (GCP)", "North America (GCP)", "Asia (excl. China and India)", "Europe (excl. EU-27)", "Europe (excl. EU-28)", "North America (excl. USA)", "South America (GCP)", "Africa (GCP)", "International shipping")

# create and plot top 10 list
top10_2023 <- annual_emmisions_data |>
  filter(Year == 2023) |>
  filter(!Entity %in% regions) |>
  filter(Entity != "World" & Entity != "Russia" & Entity != "Germany") |>
  select(Entity, `Annual CO₂ emissions`) |>
  arrange(desc(`Annual CO₂ emissions`)) |>
  slice_head(n = 10)

# plot
ggplot(top10_2023,
       aes(x = reorder(Entity, `Annual CO₂ emissions`),
           y = `Annual CO₂ emissions`,
           fill = Entity)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 10 CO2 Emitting Countries (+ World) in 2023",
    x = "Country",
    y = "Annual CO2 Emissions (tonnes)",
  ) +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  scale_fill_tableau() +
  theme_minimal()
```

## Question 2a: Which Countries Have Emitted The Most Carbon Dioxide In Total?

```{r}
countries <- c("United States", "China", "India", "Europe", "Japan")

# compute cumulative emissions by country over time
co2_countries <- annual_emmisions_data |>
  filter(Entity %in% countries) |>
  select(Entity, Year, `Annual CO₂ emissions`) |>
  group_by(Entity) |>
  arrange(Year, .by_group = TRUE) |>
  mutate(Cumulative = cumsum(`Annual CO₂ emissions`)) |>
  ungroup()

# pick milestone years for clearer bars
milestones <- c(1800, 1900, 1950, 2000, 2025)
co2_milestones <- co2_countries |>
  filter(Year %in% milestones)

# plot stacked bar chart
ggplot(co2_milestones,
       aes(x = factor(Year), y = Cumulative / 1e9, fill = Entity)) +
  geom_col() +
  labs(
    title = "Cumulative CO2 Emissions by Country (1800's–2000's)",
    x = "Time Period",
    y = "Cumulative CO2 Emissions (billion tonnes)",
    fill = "Country"
  ) +
  scale_y_continuous(labels = label_number(accuracy = 1)) +
  scale_fill_tableau() +
  theme_minimal(base_size = 12)
```

## Question 2b: Which Countries Have Emitted The Most Carbon Dioxide In Total?

```{r}
# pick the latest available year
latest_year <- max(co2_countries$Year, na.rm = TRUE)

# calculate share of global cumulative emissions
shares <- co2_countries |>
  filter(Year == latest_year) |>
  group_by(Year) |>
  mutate(share = 100 * Cumulative / sum(Cumulative)) |>
  ungroup() |>
  arrange(desc(share)) |>
  slice_head(n = 7)

# plot
ggplot(shares, 
       aes(x = reorder(Entity, share), 
           y = share, 
           fill = Entity)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = paste("Top 7 Cumulative Carbon Dioxide Emitters"),
    x = "Country",
    y = "Share of global cumulative Carbon Dioxide emissions (%)"
  ) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) +
  scale_fill_tableau() +
  theme_minimal(base_size = 13)
```

## Question 3: Which Countries Emit The Most Carbon Dioxide Per Person?

```{r}
# combine carbons emissions and population data
population_data <- read_excel("datasets/Country-Year-Population.xlsx")

emissions_2023 <- annual_emmisions_data |>
  filter(Year == 2023) |>
  filter(!Entity %in% regions) |>
  select(Code, Year, Annual_CO2_Emissions = `Annual CO₂ emissions`)

data_2023 <- population_data |>
  select(`Country Name`, `Country Code`, `2023 [YR2023]`) |>
  rename(
    Country_Name = `Country Name`,
    Country_Code = `Country Code`,
    Population_2023 = `2023 [YR2023]`
  )

combined_2023 <- emissions_2023 |>
  left_join(data_2023, by = c("Code" = "Country_Code"), relationship = "many-to-many") |>
  mutate(
    Population_2023 = as.numeric(Population_2023),
    CO2_per_capita = Annual_CO2_Emissions / Population_2023
  )
```

```{r}
countries_of_interest <- c("United States", "India", "China", "Germany","France", "Japan", "Indonesia")

selected_countries <- combined_2023 |>
  filter(Country_Name %in% countries_of_interest)

ggplot(selected_countries,
       aes(x = reorder(Country_Name, CO2_per_capita),
           y = CO2_per_capita,
           fill = Country_Name)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "CO2 Emissions per Capita (2023)",
    x = "Country",
    y = "CO2 Emissions per Capita (tonnes per person)"
  ) +
  scale_y_continuous(labels = label_number(accuracy = 0.1)) +
  scale_fill_tableau() +
  theme_minimal()
```